---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "50%",
  fig.align='right'
)
```


<!-- badges: start -->
[![Travis build status](https://travis-ci.com/softloud/neet.svg?branch=master)](https://travis-ci.com/softloud/neet)
<!-- badges: end -->

# `neet::` Test for non-empty thing of expected type

A *neet test* tests for **non-empty** thing of **expected type**. This
is what is referred to as a *boundary condition* test in RStudio's primers
(todo: citation).

These expectations are intended to integrate into minimal testing workflow
*for development* of data analyses. When developing a function, we will need parameters, and structure of the pipeline. These tests
enable the developer to feel reassured the pipeline's functions are
outputting non-empty thing of expected type, while the developer decides the best structure for an analysis pipeline. 

A `character` string will be checked for being of string-length > 1.

A `numeric` is checked for not being `NA`, `NULL`, `Inf`, or `-Inf`.

A `list`` is checked for being of length > 1.

A `data.frame` is checked for having at least one row.

# Installation

You can install the released version of `neet` from GitHub with:

``` r
# install devtools
install.packages(devtools) 

# use devtools to install package from github
devtools::install_github("softloud/neet")

```

## Example


```{r example}
library(testthat)
library(neet)

## basic example code

# assertions check inputs of code in function scripts
assert_neet(3, "numeric") # test numeric
assert_neet("cat", "character") # test character string
assert_neet(mtcars, "data.frame") # test data frame

# tests to check neet expectations in testthat files
test_neet(3, "numeric") # test numeric
test_neet("cat", "character") # test character string
test_neet(mtcars, "data.frame") # test data frame

```

# Workflow

The `neet::` package supports a coding workflow.

```{r }
workflow(0)
```

This section describes the test-driven workflow presented above as a `code::proof`ed coding to doneness.  

The model workflow comprises three stages, repeated three times, before returning to the start, the `code::registration`. We use *model* to suggest the workflow may be adapted for different use-cases.

Each phase consists of:

1. `code::registration`
2. tests
3. code

## coding to `code::proof` workflow

The tests vary each time in complexity, so that the complete model cycle consists of ten phases of work:

1. `code::registration`

```{r message=FALSE}
workflow(1)
```

In an issue on GitHub:

1. Describe the algorithm's intended purpose.
2. Describe the input parameters and how they will be tested.
3. Describe the output parameters and how they will be tested.

## purpose

A `duck` that `"quack"`s.

## input

Nothing.

## output

- [ ] one neet
- [ ] all neets
- [ ] and the rest


Character string "quack".

2. `neet` tests, one per function

```{r}
workflow(2)
```

Write a test to check the duck quacks at all. This test currently fails, because I have not written the code to make it pass, yet. 

```{r error=TRUE}
test_that(expect_match(duck(), "quack"))
```

3. code

```{r}
workflow(3)
```

Write code to make the test pass.

```{r eval=FALSE}
# write function
duck <- function() {"quack!"}
```

And now no error message is returned.

```{r error=TRUE}
test_that(expect_match(duck(), "quack"))
```



4. `code::registration`

Either update or rewrite the `code::registration`. It becomes more obvious what code needs to be written once the developer has spent some time again with the function. 

- [x] one neet
- [ ] all neets
- [ ] and the rest


5. `neet` tests, for all inputs for each function

```{r}
workflow(5)
```

Now consider all inputs. Since we have nothing, we might consider the case where there is something, and write a test that expects an error.

```{r}
test_that("function fails with input", {
  expect_error(duck(3))
})
```



6. code

The duck quacks sufficiently. No more code required! 


7. `code::registration`

```{r}
workflow(7)

```

Either update or rewrite the `code::registration`.

- [x] one neet
- [x] all neets
- [ ] and the rest


8. tests, and the rest, i.e., any other cases to test for

I am satisified with the test coverage. 

9. code

The duck quacks sufficiently.

7. `code::registration`

```{r}
workflow(10)
```

- [x] one neet
- [x] all neets
- [x] and the rest

For the final `code::registration`, update the document for your future self with details you may need when implementing new features.


