---
title: "When is done *done*? Practice fundamentals of automated testing for collaborative data analysis development"
author: "Charles T. Gray, Emily Riederer, Daniel Fryer, and Elise Gould"
bibliography: references.bib
output: 
  html_document:
    number_sections: true
# output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Key arguments to keep in mind:

- two key arguments: workflow towards doneness and also effective collaboration
- how to practice testing? test-driven development
- *not* what is best-practice
- *not* comprehensive guide to testing
- collaborative --> anxiety reducing
- in-dev testing scripts are gold
- preregistration & qrps

# Fundamentals of practice

Chuang C. Chang's _Fundamentals of Piano Practice_ sets out to address a gap in piano pedagogy [@chang_fundamentalspianopractice_2009a]. A similar gap exists in the implementation of *good enough* practices [@wilson_good_2017], which is to say, what we might reasonably expect of a analyst, in reproducible computing for research data analyses. The objective is different from advanced pianism, but we characterise testing *practice*, as opposed to *techique*, analogous to how Change delineates between  *piano* practice and technique. Through attempting to identify the fundamentals of automated testing for collaborative data analysis development, this manuscript aims to articulate the gap in understanding automated testing implementation for analysts. 

Testing as described in *R Packages* [todo: cite] and the primers provided by RStudio [todo: cite],   

As Chang notes, whilst there a rich history of technical pedagogy, there is a dearth of guidance for pianists on _learning_ the technique [@chang_fundamentalspianopractice_2009a]. There are many canonical texts of pianistic technique pedagogy. Bach provides a pathway from small canons [todo: cite], to two-part inventions[todo: cite], three-part sinfonia, and the challenge of *The Well-Tempered Clavier* and *The Art of Fugue*. Bartok provides the *Mikrokosmos* [todo: ], and Czerny's *School of Velocity* [to do]. In each case, technical exercises of increasing difficulty are provided. In piano pedagogy, a _technical_ exercise isolates a particular aspect of pianistic technique [todo examples]. For example, [todo Czerny staccato]. Or, [todo voicing technique].

The dearth that Chang attempts to address is in pianistic _practice_ habits that will lead to succesful adoption of these techniques. In science, we might call this _work flow_ [todo cite]. 

> .. practically every piano learning method consists of showing students what to practice, and what kinds of techniques (runs, arpeggios, legato, staccato, trills, etc.) are needed. There are few instructions on how to practice order  to be able to play them, which is mostly left to the student and endless repetitions [@chang_fundamentalspianopractice_2009a].

Wilson et al. followed their work on _best practices_  in scientific computing [@wilson_best_2014], with a reflection on _good enough_ practices [@wilson_good_2017], in recognition that we must account for what we might reasonably request of practitioners of data analysis. In this manuscript, we consider one component of *good enough* practice in data analysis: *automated testing*. 

Automated tests are a formalised way of implementing checks that inputs and outputs of algorithsms are as expected [@wickham_r_2015]. Informative messages are output that assist the developer in identifying where code is not performing as expected. 

## Collaboration via automated testing 

At heart, automated tests are collaborative. This may be with others, but applies at least as much to a analyst approaching their own past work with which they have become unfamiliar. Automated tests provide an efficient way of returning to and extending an analysis. 

Reproducible research compendia provide a means by which analysts can share their work so that it may be extended by others, and automated tests provide `code::proof` [@grayCodeProofPrepare2019], a measure of confidence in the algorithm for others. Hayes expresses concern about using algorithms published without automated tests [@hayes_testing_2019]. However, only one quarter of the largest repository of R packages, [The Comprehensive R Archive Network](https://cran.r-project.org/), have any automated tests [@hester_covr_2016], highlighting that, despite testing identified as a 'vital' part of data analysis [@wickham_r_2015], automated testing is yet to be widely adopted.

Indeed, as noted in Wilson's testing primers (in development) for [RStudio](https://rstudio.cloud/learn/primers) [@_rstudiocloud_], 

> Almost all (92%) of the catastrophic system failures are the result of incorrect handling of non-fatal errors explicitly signaled in software. In 58% of the catastrophic failures, the underlying faults could easily have been detected through simple testing of error handling code [@yuan_simpletestingcan_2014].

Scientific errors that are inadvertently introduced by faulty workflow are characterised in metascience as *questionable research practices*.

## Questionable research practices in scientific computing

One way to characterise automated tests is to consider them part of a suite of methodologies that are being developed to mitigate _questionable research practices_ [@fraser_questionable_2018]. For social scientists and ecologists, a literature is emerging on how our methodologies inadvertently lead to errors in scientific reasoning, $p$-hacking, altering a statistical model to achieve a desired outcome. But there are others, too, *HARKing*, hypothesising after the fact, where the question is changed to achieve a desired outcome from the model, and *cherry picking*, leaving out troublesome aspects of the analysis that don't fit a desired narrative. 

Gould is extending these ideas to an examination of the workflow of ecological models, adapting Gelman and [todo]  

[todo: gould extension bridge to computing]

As researcher software engineers, it behoves us to consider what are questionable research practices in software produced for data analyses. Version control and open code sharing via a platform such as GitHub, is one way to mitigate questionable research practices in scientific computing [@Bryan2017ExcuseMD]. There is also a growing literature on reproducible research compendia via packaged analyses [@marwick_packaging_2018; @wilson_good_2017].     

This manuscript contributes to this literature by focussing on workflows for automated tests that assist the developer communicate what they have coded to others and their future self, as a means of mitigating questionable research practices in scientific computing. 

# Test-driven development

## The benefits of preregistring code

> inherently collaborative

# A test-driven toolchain walkthrough for an in-development analysis

In this section we now consider the practicality of implementing tests through a *toolchain walkthrough*, an opinionated documetation of a scientific workflow , towards a measure of `code::proof`, confidence in the algorithm implemented [@grayCodeProofPrepare2019]. In particular, a toolchain walkthrough is a reflection of *one* workflow, whilst others, indeed, better, workflows might exist. This is in contrast to a comprehensive review of tools. Instead, a toolchain walkthrough ruminates on considerations *after* a set of tools have been chosen.

Toolchain walkthroughs aim to identify obstacles and advantages in implementation of scientific workflows. By necessity, a toolchain walkthrough is language specific, but, ideally, observations emerge that are useful for any appropriate language employed for data analysis, such as Python. This toolchain walkthrough examines the *process*, analogous to piano *practice*, of implementing tests, as opposed to defining comprehensively the nature of *good enough* tests, analogous to guidance on pianistic technique.

## Objective of this toolchain walkthrough

This toolchain walkthrough aims to provide guidance on implementing a test-driven workflow for an in-development analysis. Many analyses begin as scripts that develop [todo expan] [@riederer_rmarkdowndrivendevelopment_2019]. The central question of this manuscript is what constitutes a minimal level of testing for in-development analysis, where code is still being written and features implemented. Automated tests assist in time-consuming debugging tasks [@wickham_r_2015], but also in providing information with a developer who is unfamiliar with the code. 

This is a first effort in identifying the fundamentals of automated testing for the collaborative process of developing an analysis in R. Analgous to Riederer's *RMarkdown-driven development* [@riederer_rmarkdowndrivendevelopment_2019], which deconstructs the workflow of developing an analysis from .Rmd notebook-style reports to packaged analyses, we consider a set a computational tools that form a workflow to assist in the coherent development of automated tests for data analysis. This is an extension of the workflow suggestions provided in *R Packages*, with a specific focus on collaborative workflows in research.   

## Devops and assumed expertise

This toolchain walkthrough assumes a knowledge of R advanced enough to be using the language to be answering scientific research claims.

- tools used: testthat, neet, covr
- GitHub


## Get an overview with `covr::`

With a packaged analysis, it's easy to get a sense of existing automated testing coverage, if any tests have been written. The analysis functions provided by `varameta::` were put on hold while signficant discussion was written.   

``` r
library(covr)
package_coverage("~/Documents/repos/varameta/")
#> varameta Coverage: 70.71%
#> R/dist_name.R: 0.00%
#> R/g_cauchy.R: 44.44%
#> R/g_norm.R: 71.43%
#> R/hozo_se.R: 92.31%
#> R/bland_mean.R: 100.00%
#> R/bland_se.R: 100.00%
#> R/effect_se.R: 100.00%
#> R/g_exp.R: 100.00%
#> R/g_lnorm.R: 100.00%
#> R/hozo_mean.R: 100.00%
#> R/wan_mean_C1.R: 100.00%
#> R/wan_mean_C2.R: 100.00%
#> R/wan_mean_C3.R: 100.00%
#> R/wan_se_C1.R: 100.00%
#> R/wan_se_C2.R: 100.00%
#> R/wan_se_C3.R: 100.00%
```

<sup>Created on 2020-02-01 by the [reprex package](https://reprex.tidyverse.org) (v0.3.0)</sup>

The `covr::` function `package_coverage` provides the percentage of the lines of code run in tests. 

# A toolchain walkthrough for using automated tests `varameta::`

## Preregistering an algorithm

The Center for Open Science recommend preregistering research hypotheses before testing them as one safeguard against inadvertent questionable research practices. Analogously, we might think of preregistering our code as providing some `code::proof` of the implementation of our algorithm.



To preregister a function:

### On a piece of paper.

1. Draw a diagram of the inputs and outputs of a function.
2. Draw where this function fits in the pipeline of the package, if appropriate.

### In an issue on GitHub:

1. Describe the function's intended purpose.
2. Describe as explicitly as possible the input parameters and how they will be tested.
3. Describe as explicitly as possible the output parameters and how they will be tested.
4. Include image of diagram if not burdonsomely time consuming (but this often is).

- ggplot picture of function? use vis thingy from drake?



```{r message=FALSE, fig.width=15, echo=FALSE}
library(tidyverse)

library(ggraph)
library(tidygraph)

nodes <- tribble(
  ~node, ~state,
  "a", "input",
  "b", "input",
  "q1", "input",
  "q2", "input",
    "m", "input",
  "n","input",
  "hozo", "estimator",
  "pren_c3", "estimator",
  "pren_c1", "estimator",
  "bland", "estimator",
  "wan_c1", "estimator",
  "wan_c2", "estimator",
  "wan_c3", "estimator",
  "se_median", "output",
  "mean", "output",
  "sd", "output"
) %>% 
  mutate(state = fct_relevel(state, "input"))

edges <- tribble(
  ~from, ~to,
  # hozo estimator
  5,7, 
  6,7,
  1,7,
  2,7,
  7,15,
  7,16,

  # wan1 estimator
5,11, 
6,11,
1,11,
2,11,
11,15,
11,16,
  
  # pren1 estimator
5,9, 
6,9,
1,9,
2,9,
9,14,
  
  # bland estimator
  5,10,
  6,10,
  1,10,
  2,10,
  3,10,
  4,10,
  10,15,
  10,16,
  
# wan2 estimator
5,12,
6,12,
1,12,
2,12,
3,12,
4,12,
12,15,
12,16,

  # wan3 estimator
5,13, 
6,13,
3,13,
4,13,
13,15,
13,16,

# pren3
5,8, 
6,8,
3,8,
4,8,
8,14
)

graph <- tbl_graph(nodes, edges)

graph %>% 
  ggraph() +
  geom_edge_link(arrow = arrow(20), colour = "lightgrey") + 
  geom_node_label(aes(label = node, colour = state),  
                  size = 6,
                  fill = "lightgrey",
                  alpha = 0.6) +
  theme_graph() +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none")
  

```

```{r message=FALSE}
library(igraph)

mygraph <- make_graph(~ 
a-+hozo, b-+hozo, m-+hozo, n-+hozo, hozo-+s, hozo-+mean,
a-+wan1, b-+wan1, m-+wan1, n-+wan1, wan1-+s, wan1-+mean,
a-+pren, b-+pren, m-+pren, n-+pren, pren-+s,
q1-+wan3, q3-+wan3, m-+wan3, n-+wan3, wan3-+s, wan3-+s,
q1-+pren, q3-+pren, m-+pren, n-+pren, pren-+s,
a-+wan2, b-+wan2, q1-+wan2, q3-+wan2, m-+wan2, n-+wan2, wan2-+s, wan2-+mean,
a-+bland, b-+bland, q1-+bland, q3-+bland, m-+bland, n-+bland, bland-+s, bland-+mean
           ) 

mygraph %>% plot(
             vertex.color = "grey",
             vertex.frame.color = "darkgrey",
             vertex.size = 10,
             vertex.label.cex = 0.8,
             edge.arrow.size = 0.4
             )


```



## `neet` test the inputs and outputs

## `neet` test all input cases

##  complete tests for 


### identifying the testable components


## Workflow for in-development analysis

- If not neet than neet. If neet, then what?

# Collaboration

# Assessing the doneness of software 

# Further directions

# References
